---
title: "K-means-clustering-project"
author: "Umran YAMAN"
date: "2023-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Documents/Data_ML_projects/Data_Machine/R-Course-HTML-Notes/R-for-Data-Science-and-Machine-Learning/Training Exercises/Machine Learning Projects/CSV files for ML Projects")
```


### K Means Clustering Project
Usually when dealing with an unsupervised learning problem, its difficult to get a good measure of how well the model performed. For this project, we will use data from the UCI archive based off of red and white wines (this is a very commonly used data set in ML).
We will then add a label to the a combined data set, we'll bring this label back later to see how well we can cluster the wine into groups.
Get the Data
Download the two data csv files from the UCI repository (or just use the downloaded csv files).
Use read.csv to open both data sets and set them as df1 and df2. Pay attention to what the separator (sep) is.

```{r}
df1 <- read.csv("winequality-red.csv", sep = ";")
df2 <- read.csv("winequality-white.csv", sep = ";")
```

Now add a label column to both df1 and df2 indicating a label 'red' or 'white'.

```{r}
df1$label <- "red"
df2$label <- "white"
```

Check the head of df1 and df2.

```{r}
head(df1)
head(df2)
```
Combine df1 and df2 into a single data frame called wine.

```{r}
wine <- rbind(df1, df2)
str(wine)
```
### EDA
Let's explore the data a bit and practice our ggplot2 skills!
Create a Histogram of residual sugar from the wine data. Color by red and white wines.

```{r}
ggplot(wine, aes(residual.sugar)) + geom_histogram(aes(fill=label), color = 'black', bins=50, alpha = 0.7) + scale_fill_manual(values = c("darkred", "beige")) + theme_bw()
```

Create a Histogram of citric.acid from the wine data. Color by red and white wines.

```{r}
ggplot(wine, aes(citric.acid)) + geom_histogram(aes(fill=label), color = 'black', bins=50, alpha = 0.7) + scale_fill_manual(values = c("darkred","beige")) + theme_bw()
```
Create a Histogram of alcohol from the wine data. Color by red and white wines.


```{r}
ggplot(wine, aes(alcohol)) + geom_histogram(aes(fill=label), color = 'black', bins=50, alpha = 0.7) + scale_fill_manual(values = c("darkred","beige")) + theme_bw()
```


```{r}
#Create a scatterplot of residual.sugar versus citric.acid, color by red and white wine.
ggplot(wine, aes(citric.acid, residual.sugar)) + geom_point(aes(color=label), alpha=0.2) + scale_color_manual(values=c("darkred", "beige")) + theme_dark()

```
There is so much overlapping and noise. 

Create a scatterplot of volatile.acidity versus residual.sugar, color by red and white wine.
```{r}
ggplot(wine, aes(volatile.acidity, residual.sugar)) + geom_point(aes(color=label),alpha=0.2) + scale_color_manual(values=c("darkred", "beige")) + theme_dark()
```
- volatile.acidity may help us to cluster, probably rose or other types are inbetween the classes.

Grab the wine data without the label and call it clus.data

```{r}
clus.data <- wine[,1:12]
head(clus.data)
```

### Building the Clusters

Call the kmeans function on clus.data and assign the results to wine.cluster.

```{r}
set.seed(101)
wine.cluster <- kmeans(clus.data, 2)

#center values of each features
print(wine.cluster$centers)
```
### Evaluating the Clusters

You usually won't have the luxury of labeled data with KMeans, but let's go ahead and see how we did!
Use the table() function to compare your cluster results to the real results. Which is easier to correctly group, red or white wines?

```{r}
#this is not a confusion matrix, we are not doing a classificaiton, it is clustering
table(wine$label, wine.cluster$cluster)
```
There seems to be noise in white wines, maybe due to rose being classified as white wine, or vice versa. Overall this makes sense, the chemical features may not be very helpful for the clustering. Domain knowledge comes in place. 
