---
title: "neural nets"
author: "Umran YAMAN"
date: "2023-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Neural Net Project
Let's wrap up this course by taking a a quick look at the effectiveness of Neural Nets!
We'll use the Bank Authentication Data Set from the UCI repository.
The data consists of 5 columns:
variance of Wavelet Transformed image (continuous)
skewness of Wavelet Transformed image (continuous)
curtosis of Wavelet Transformed image (continuous)
entropy of image (continuous)
class (integer)
Where class indicates whether or not a Bank Note was authentic.

Get the Data
Use read.csv to read the bank_note_data.csv file.

```{r}
setwd("~/Documents/Data_ML_projects/Data_Machine/R-Course-HTML-Notes/R-for-Data-Science-and-Machine-Learning/Training Exercises/Machine Learning Projects/CSV files for ML Projects")
bank <- read.csv('bank_note_data.csv')
```

Check the head of the data frame and its structure.

```{r}
head(bank)
```

EDA
Create whatever visualizations you are interested in. We'll skip this step for the solutions notebook/video because the data isn't easily interpretable since its just statistical info on images.
Train Test Split
Use the caTools library to split the data into training and testing sets.

```{r}
library(caTools)
set.seed(101)
sample <- sample.split(bank$Class, SplitRatio = 0.7)
train <- subset(bank, sample == T)
test <- subset(bank, sample==F)
```

Check the structure of the train data and note that Class is still an int data type. We won't convert it to a factor for now because the neural net requires all numeric information.

```{r}
str(train)
```

Building the Neural Net
Call the neuralnet library

```{r}
library(neuralnet)
```

Browse through the documentation of neuralnet

```{r}
help("neuralnet")
```

Use the neuralnet function to train a neural net, set linear.output=FALSe and choose 10 hidden neurons (hidden=10)

```{r}
#hidden neurons number is tricky, we will do it as in last time
nn <- neuralnet(Class ~ Image.Var + Image.Skew + Image.Curt + Entropy, data = train, linear.output = FALSE, hidden = c(5,3))


#Predictions
#Use compute() to grab predictions useing your nn model on the test set. Reference the lecture on how to do this.
predicted.nn.values <- compute(nn, test[1:4])
```


```{r}
head(predicted.nn.values$net.result)
```

```{r}
predictions <- sapply( predicted.nn.values$net.result, round)
head(predictions)
```

```{r}
#to compare
library(randomForest)
library(caTools)
set.seed(101)
bank$Class <- factor(bank$Class)
sample <- sample.split(bank$Class, SplitRatio = 0.7)
train <- subset(bank, sample == T)
test <- subset(bank, sample==F)

rf.model <- randomForest(Class ~ ., data = train)

rf.pred <- predict(rf.model, test)
table(rf.pred, test$Class)
#reality check. 
```

