---
title: "Linear_Regression_Project"
author: "Umran YAMAN"
date: "2023-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Regression Project

# Get the Data

You can download the data or just use the supplied csv in the repository. The data has the following features:
datetime - hourly date + timestamp
season - 1 = spring, 2 = summer, 3 = fall, 4 = winter
holiday - whether the day is considered a holiday
workingday - whether the day is neither a weekend nor holiday
weather -
1: Clear, Few clouds, Partly cloudy, Partly cloudy
2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
temp - temperature in Celsius
atemp - "feels like" temperature in Celsius
humidity - relative humidity
windspeed - wind speed
casual - number of non-registered user rentals initiated
registered - number of registered user rentals initiated
count - number of total rentals

```{r}
setwd("~/Documents/Data_ML_projects/Data_Machine/R-Course-HTML-Notes/R-for-Data-Science-and-Machine-Learning/Training Exercises/Machine Learning Projects/CSV files for ML Projects")
df <- read.csv("bikeshare.csv")
```

```{r}
head(df)
```
## Exploratory Data Analysis

Create a scatter plot of count vs temp. Set a good alpha value.

```{r}
library(ggplot2)
ggplot(df, aes(x=temp, y=count)) + geom_point(alpha=0.3, aes(color=temp)) +theme_bw()
```

Plot count versus datetime as a scatterplot with a color gradient based on temperature. You'll need to convert the datetime column into POSIXct before plotting.

```{r}
df$datetime = as.POSIXct(df$datetime)
ggplot(df, aes(datetime, count)) + geom_point(aes(color = temp), alpha = 0.5) + scale_color_continuous(low='#55D8CE', high='#FF6E2E') +theme_bw()
``` 
Pros and cons right now of Linear Regression:

Pros:
Simple to explain
Highly interpretable
Model training and prediction are fast
No tuning is required (excluding regularization)
Features don't need scaling
Can perform well with a small number of observations
Well-understood

Cons:
Assumes a linear relationship between the features and the response
Performance is (generally) not competitive with the best supervised learning methods due to high bias
Can't automatically learn feature interactions


What is the correlation between temp and count?
```{r}
cor.data = cor(df[, c('temp', 'count')])
cor.data
```

```{r}
ggplot(df, aes(factor(season), count)) + geom_boxplot(aes(color=factor(season))) +theme_bw()
```

The most favored season is 2 and 3. The bike rental counts are showing seasonality. 

Notice what this says:
A line can't capture a non-linear relationship.
There are more rentals in winter than in spring

We know of these issues because of the growth of rental count, this isn't due to the actual season!

### Feature Engineering

A lot of times you'll need to use domain knowledge and experience to engineer and create new features. Let's go ahead and engineer some new features from the datetime column.
Create an "hour" column that takes the hour from the datetime column. You'll probably need to apply some function to the entire datetime column and reassign it. Hint:
time.stamp <- bike$datetime[4]
format(time.stamp, "%H")

```{r}
time.stamp <- df$datetime[4]
format(time.stamp, "%H")


df$hour = format(df$datetime, "%H")
head(df)
```

Now create a scatterplot of count versus hour, with color scale based on temp. Only use bike data where workingday==1.
Optional Additions:
Use the additional layer: scale_color_gradientn(colors=c('color1',color2,etc..)) where the colors argument is a vector gradient of colors you choose, not just high and low.
Use position=position_jitter(w=1, h=0) inside of geom_point() and check out what it does.

```{r}
df_subsetted = subset(df, workingday==1)

#jitter fills out the data points

ggplot(df_subsetted, aes(x=hour, y=count)) + geom_point(aes(color=temp),position=position_jitter(w=1, h=0)) + scale_color_gradientn(colors = c("dark blue", "blue", "lightblue", "lightgreen", "yellow", "orange", "red")) 

df_subsetted_none = subset(df, workingday==0)
ggplot(df_subsetted_none, aes(x=hour, y=count)) + geom_point(aes(color=temp),position=position_jitter(w=1, h=0)) + scale_color_gradientn(colors = c("dark blue", "blue", "lightblue", "lightgreen", "yellow", "orange", "red")) 
```

You should have noticed that working days have peak activity during the morning (~8am) and right after work gets out (~5pm), with some lunchtime activity. While the non-work days have a steady rise and fall for the afternoon
Now let's continue by trying to build a model, we'll begin by just looking at a single feature.
Building the Model
Use lm() to build a model that predicts count based solely on the temp feature, name it temp.model

```{r}
#The data has the seasonality, the linear regression wont be that helpful, "time series" version of the data does not enable the random choosing. Thats why we are not splitting 

temp.model = lm(count ~ temp, df)
summary(temp.model)
```
How many bike rentals would we predict if the temperature was 25 degrees Celsius? Calculate this two ways:
Using the values we just got above
Using the predict() function
You should get around 235.3 bikes.

```{r}
temp.predictions = predict(temp.model)
```

```{r}
results <- cbind(temp.predictions, df$temp)
colnames(results) <- c('predicted', 'actual')
results <- as.data.frame(results)
print(head(results))


# How many bike rental counts at 25C?
6.0462 + 9.17 * 25

df$hour = sapply(df$hour, as.numeric)
class(df$hour)
```

Finally build a model that attempts to predict count based off of the following features. Figure out if theres a way to not have to pass/write all these variables into the lm() function. Hint: StackOverflow or Google may be quicker than the documentation.

season
holiday
workingday
weather
temp
humidity
windspeed
hour (factor)

```{r}
all.model = lm(count ~ . -casual -registered -datetime -atemp, df)
summary(all.model)
```

