---
title: "Tree Methods Project"
author: "Umran YAMAN"
date: "2023-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Tree Methods Project
For this project we will be exploring the use of tree methods to classify schools as Private or Public based off their features.
Let's start by getting the data which is included in the ISLR library, the College data frame.
A data frame with 777 observations on the following 18 variables.
Private A factor with levels No and Yes indicating private or public university
Apps Number of applications received
Accept Number of applications accepted
Enroll Number of new students enrolled
Top10perc Pct. new students from top 10% of H.S. class
Top25perc Pct. new students from top 25% of H.S. class
F.Undergrad Number of fulltime undergraduates
P.Undergrad Number of parttime undergraduates
Outstate Out-of-state tuition
Room.Board Room and board costs
Books Estimated book costs
Personal Estimated personal spending
PhD Pct. of faculty with Ph.D.’s
Terminal Pct. of faculty with terminal degree
S.F.Ratio Student/faculty ratio
perc.alumni Pct. alumni who donate
Expend Instructional expenditure per student
Grad.Rate Graduation rate

```{r}
library(ISLR)
head(College)
```

Get the Data
Call the ISLR library and check the head of College (a built-in data frame with ISLR, use data() to check this.) Then reassign College to a dataframe called df

```{r}
df <- College
```


EDA
Let's explore the data!
Create a scatterplot of Grad.Rate versus Room.Board, colored by the Private column.

```{r}
library(ggplot2)
ggplot(df, aes(Room.Board, Grad.Rate)) + geom_point(aes(color = Private),alpha=0.5)
#Create a histogram of full time undergrad students, color by Private.
ggplot(df, aes(F.Undergrad)) + geom_histogram(aes(fill = Private), color='black', bins = 50, alpha=0.5) +theme_bw()
```
```{r}
#Create a histogram of Grad.Rate colored by Private. You should see something odd here.
ggplot(df, aes(Grad.Rate)) + geom_histogram(aes(fill = Private), color='black', bins =50, alpha=0.5) + theme_bw()
```
What college had a Graduation Rate of above 100% ?

-Private college , Cazenovia College

Change that college's grad rate to 100%

```{r}
subset(df, Grad.Rate >100)
df["Cazenovia College", "Grad.Rate"] <- 100
df["Cazenovia College", "Grad.Rate"]
```
Train Test Split
Split your data into training and testing sets 70/30. Use the caTools library to do this.

```{r}
library(caTools)
set.seed(101)
sample <- sample.split(df$Private, SplitRatio = 0.7)
train <- subset(df, sample==T)
test <- subset(df, sample==F)
```

Decision Tree
Use the rpart library to build a decision tree to predict whether or not a school is Private. Remember to only build your tree off the training data.

```{r}
library(rpart)
#Train model
tree <- rpart(Private ~ . , method='class', data= df)
printcp(tree)
```
```{r}
#Use predict() to predict the Private label on the test data.

tree.pred <- predict(tree, test)
head(tree.pred)
#Probabilities of model yes or no about being private or not
```
```{r}
#Change two columsn to one column
tree.preds <- as.data.frame(tree.pred)

joiner <- function(x){
  if (x >= 0.5){
    return('Yes')
  }else{
    return('No')
  }
}

tree.preds$Private <- sapply(tree.preds$Yes, joiner)
head(tree.preds)
#We created a Private Column to see transform the probablilites to yes or no answer 
```
```{r}
#Confusion matrix
table(tree.preds$Private, test$Private)
```

```{r}
#Plot the decision tree
library(rpart.plot)
prp(tree)
```

Random Forest
Now let's build out a random forest model!
Call the randomForest package library

```{r}
library(randomForest)
#What was your model's confusion matrix on its own training set? Use model$confusion.

model <- randomForest(Private ~ .,   data=train, importance=TRUE)
print(model)
model$confusion
```


Grab the feature importance with model$importance. Refer to the reading for more info on what Gini[1] means.[2]


```{r}
model$importance
```

Predictions
Now use your random forest model to predict on your test set!

```{r}
model.pred <- predict(model, test)
table(model.pred, test$Private)
```
It should have performed better than just a single tree, how much better depends on whether you are emasuring recall, precision, or accuracy as the most important measure of the model.
